
\documentclass{article}

\usepackage{amstext}
\usepackage{indentfirst}
\usepackage{url}

\newcommand{\set}[1]{\{\, #1 \,\}}

\begin{document}

\title{Rao Tests}

\author{Charles J. Geyer}

\maketitle

\section{Introduction}

Let $l_n$ be the log likelihood for sample size $n$ for some statistical
model, and let $\theta^*_n$ be the maximum likelihood estimate (MLE)
for a smooth submodel.
The Rao test statistic is
\begin{equation} \label{eq:rao}
   R_n = 
   \bigl(\nabla l_n(\theta^*_n)\bigr)^T J_n(\theta^*_n)^{-1}
   \nabla l_n(\theta^*_n),
\end{equation}
where 
$$
   J_n(\theta) = - \nabla^2 l_n(\theta)
$$
is the observed Fisher information matrix for sample size $n$
(this is copied from the lecture notes
\url{http://www.stat.umn.edu/geyer/8112/notes/tests.pdf}).

Note that $l_n$ and $J_n$ are for the larger model of the two nested
models being compared and that $\theta^*_n$ is the MLE for the smaller
model of the two \emph{expressed in the parameter space of the larger
model}.

\section{Implementation}

This is part of the implementation of method \texttt{glmdr} of R generic
function \texttt{anova}.  Part of this function checks whether models
are nested (see design document \texttt{nested.pdf} in this directory).
Part of this function does LRT (likelihood ratio tests,
a.~k.~a.\ analysis of deviance).  This document does not cover that either.

As the introduction says, we need three things.
\begin{itemize}
\item The MLE for the smaller model, expressed in the parameter space of
the larger model (it already having being checked that the former is a submodel
of (is nested within) the latter).   Since the check for nesting is completely
general, this step is nontrivial.
\item The first derivative vector for the larger model evaluated at the MLE
for the smaller model calculated in step one.
\item The second derivative matrix for the larger model evaluated at the MLE
for the smaller model calculated in step one.
\end{itemize}

\subsection{MLE}

Let $\hat{\beta}_\text{small}$ denote the MLE in the parameterization
of the smaller model.  Let $a_\text{small}$ and $a_\text{large}$ be the
offset vectors of the smaller and larger models and
$M_\text{small}$ and $M_\text{large}$ the model matrices.

The the problem of step one in our list is to find a $\hat{\beta}_\text{large}$
that satisfies
$$
   a_\text{small} + M_\text{small} \hat{\beta}_\text{small}
   =
   a_\text{large} + M_\text{large} \hat{\beta}_\text{large}
$$
Hence $\hat{\beta}_\text{large}$ is any solution for $\beta$ of the
linear system
$$
   M_\text{large} \beta
   =
   a_\text{small} - a_\text{large} + M_\text{small} \hat{\beta}_\text{small}
$$
(if the solution is not unique, then nonuniqueness does not matter).

As usual in R, we use R functions \texttt{qr} and \texttt{qr.solve} to
solve linear equations.
\begin{verbatim}
foo <- offset.small - offset.large +
    modmat.small %*% beta.hat.small
bar <- qr(modmat.large)
beta.hat.large <- qr.solve(bar, beta.hat.small)
\end{verbatim}

\subsection{Derivatives}

There is no accessible function to find derivatives of the models R package
\texttt{glmdr} fits.  But there is a non-accessible function that we can
use inside code of the package.  This is the R function returned by
R function \texttt{make.mlogl} defined in the file \texttt{R/glmdr.R}
(R function \texttt{make.mlogl} is a higher-order function that returns
function).  An example of the use of this function is inside R function
\texttt{glmdr}.

We would do something like
\begin{verbatim}
mlogl <- make.mlogl(modmat.large, resp, offset.large, family)
mout <- mlog(beta.large)
\end{verbatim}
where \texttt{beta.large} is as discussed above.

Now the Rao test statistic is
\begin{verbatim}
as.numeric(t(mout$gradient) %*% mout$hessian %*% mout$gradient)
\end{verbatim}

\end{document}

